---
title: "Two Dimensional Box and Error Plots"
author: "Jason Bryer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{boxplot2d}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ggBoxplot2d)
```

## Two Dimensional Box and Whisker Plots

Basic example using the `mtcars` dataset.

```{r boxplot2d_mtcars}
library(ggBoxplot2d)
data(mtcars)
p <- ggplot(mtcars, aes(x = wt, y = mpg, group = factor(cyl), 
						color = factor(cyl), fill = factor(cyl))) +
	geom_boxplot2d() +
	theme(legend.position = 'bottom')
p
```

We can add marginal density plots using the `ggExtra` package. We have to add the raw data points for the marginal plots to display. By setting `alpha = 0` the points are present for the density plots but are hidden from the plot.

```{r boxplto2d_ggextra}
ggExtra::ggMarginal(p + geom_point(alpha = 0), 
					type = 'density', 
					groupColour = TRUE, 
					groupFill = TRUE)
```


## Example with Palmer's Penguins

```{r}
library(palmerpenguins)
data(penguins, package = 'palmerpenguins')

ggplot(penguins, aes(x = flipper_length_mm, y = bill_length_mm, color = species, fill = species)) +
	geom_boxplot2d()
```



# References

Persaud, N., & Yates, B. T. (complete 9-chapter book manuscript submitted for hopefully final review 2022 March 25).  Cost-inclusive evaluation: Planning it, doing it, using it. Book manuscript already under contract. New York, NY: Guilford Press.

Tukey, John W. (1977). *Exploratory Data Analysis.* Addison-Wesley.

Wickham. H. (2016). *ggplot2: Elegant Graphics for Data Analysis.* Springer-Verlag


